---
title: "Final Project Draft"
author: "Anwesha Guha"
date: "2/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(fivethirtyeight)
library(fivethirtyeightdata)
library(colorblindr)

all_college <- college_all_ages
grad <- college_grad_students
recent_grad <- college_recent_grads

hate_crimes <- hate_crimes

```

## About the Data

Data was taken from the publicly available *fivethirtyeightdata* housed on GitHub. Specifically, I used the grad-students and recent-grads data, which are used in the “Economic Guide to Picking a College Major” [article](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/). The data is originally from the 2010-2012 American Community Survey and generally discusses earnings, graduate student information, and labor force outcomes.

## Plot 1
```{r initial}

recent_grad$major_collapsed <- fct_collapse(recent_grad$major_category,
                                "Arts & Humanities" = c("Arts", "Interdisciplinary", "Communications & Journalism", "Humanities & Liberal Arts"),
                                "Physical & Life Science" = c("Biology & Life Science", "Physical Sciences"),
                                "Psychology & Social Science" = c("Social Science", "Psychology & Social Work"),
                                "Industry, Public Service, & Law" = c("Industrial Arts & Consumer Services", "Law & Public Policy"))

recent_grad %>% 
  ggplot(aes(sharewomen, median)) +
  geom_point(aes(color = major_collapsed), alpha = 0.7) +
  facet_wrap(~major_collapsed) +
  theme_minimal() +
  scale_color_viridis_d(option = "E") +
  theme(legend.position = "none") +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "Share of Women",
       y = "Median Income for each Major",
       title = "Representation and Pay for Women Across Majors")

# maybe include vertical line for average share of women, horizontal line for average income
```

```{r draft two}

unique(recent_grad$major_category)

recent_grad %>% 
  ggplot(aes(sharewomen, median)) +
  geom_point(aes(color = major_collapsed, shape = major_collapsed), alpha = 0.95, size = 2) +
  geom_smooth(linetype = 0, alpha = 0.2) +
#  geom_line(stat = "smooth", alpha = 0.6, linetype = "dashed", weight = 0.7, color = "gray30") +
  scale_shape_manual(values = c(0:9)) +
  theme_minimal() +
  scale_color_brewer(palette = "Set3") +
  theme(legend.position = "top",
        legend.title = element_blank()) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "Share of Women",
       y = "Median Income for each Major",
       title = "Representation and Pay for Women Across Majors",
       subtitle = "Graphs represent the median income for each Major by share of women, grouped by major category.")

?geom_line
```



## Plot 2
```{r}
recent_grad$employ_total <- recent_grad$employed_fulltime + recent_grad$employed_parttime + recent_grad$unemployed

recent_grad_prop <- recent_grad %>% 
  group_by(major_category) %>% 
  summarize(fulltime = sum(employed_fulltime)/sum(employ_total),
         parttime = sum(employed_parttime)/sum(employ_total),
         unemploy = sum(unemployed)/sum(employ_total)) %>% 
  select(major_category, fulltime, parttime, unemploy) %>% 
  pivot_longer(-major_category,
               names_to = "employment_type",
               values_to = "proportion")

recent_grad_prop %>% 
  ggplot(aes(major_category, proportion)) +
  geom_col(aes(fill = employment_type)) +
#  geom_text(aes(label=scales::percent(proportion, 1)), size = 2, hjust = 2, color = "gray30") +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(name = "Employment Type", labels = c("Full-time", "Part-time", "Unemployed")) +
  labs(y = "Proportion",
       x = NULL,
       title = "Employment Distribution for Each Graduated Major") +
  theme(legend.position = "bottom") +
  scale_y_continuous(expand = c(0,0))

# Feedback on whether geom_text() adds anything here? I felt it looked a bit too cluttered
```

```{r take two}
recent_grad_prop %>% 
  ggplot(aes(major_category, proportion, fill = employment_type)) +
  geom_col(aes(fill = employment_type)) +
  geom_text(aes(label=scales::percent(proportion, 1)), size = 3, position = position_stack(vjust = .8), color = "gray30") +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(name = "Employment Type", labels = c("Full-time", "Part-time", "Unemployed")) +
  labs(y = "Percent",
       x = NULL,
       title = "Employment Distribution for Each Graduated Major") +
  theme(legend.position = "bottom") +
  scale_y_continuous(expand = c(0,0), labels = scales::percent)

# Feedback on whether geom_text() adds anything here? I felt it looked a bit too cluttered
```

```{r take three}
recent_grad_prop %>% 
  ggplot(aes(major_category, proportion, fill = employment_type)) +
  geom_col(aes(fill = employment_type)) +
  geom_text(aes(label=scales::percent(proportion, 1)), size = 3, position = position_stack(vjust = .8), color = "gray30") +
  coord_flip()  +
  scale_y_continuous(expand = c(0,0), labels = scales::percent) +
#  scale_x_discrete(breaks = NULL) +
  theme_minimal() +
  scale_fill_OkabeIto(name = "Employment Type", labels = c("Full-time", "Part-time", "Unemployed"),  alpha = .4) +
  labs(y = "Percent",
       x = NULL,
       title = "Employment Distribution for Each Graduated Major") +
  theme(legend.position = "bottom",
        panel.grid.minor.x = element_blank())

# Feedback on whether geom_text() adds anything here? I felt it looked a bit too cluttered
```


## Plot 3
```{r}
recent_grad_sex <- recent_grad %>% 
  group_by(major_category) %>% 
  summarize(share_men = sum(men)/sum(total),
            share_women = sum(women)/sum(total),
            salary = mean(median)) %>% 
  select(major_category, share_men, share_women, salary) %>% 
  pivot_longer(-c(major_category, salary),
               names_to = "sex",
               values_to = "share")


recent_grad_sex %>% 
  filter(major_category != "Agriculture & Natural Resources") %>% 
  ggplot(aes(reorder(major_category, -share), share)) +
  geom_col(aes(fill = sex), position = "dodge") +
#  geom_text(aes(label=scales::percent(share, 1)), size = 2, hjust = 2, position = position_dodge(0.9)) +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(name = "Sex (Reported)", labels = c("Men", "Women"), palette = "Accent") +
  labs(y = "Share",
       x = NULL,
       title = "Gender Distribution for Each Graduated Major") +
  theme(legend.position = "bottom") +
  scale_y_continuous(expand = c(0,0))

# I would be interested to also include the median salary for each field, but am stuck on how to make that work well visually. Also can't get geom_text() to work here; feedback would be great!

```


```{r take two}

recent_grad_sex %>% 
  filter(major_category != "Agriculture & Natural Resources") %>% 
  ggplot(aes(reorder(major_category, -share), share)) +
  geom_col(aes(fill = sex)) +
  geom_text(aes(label=scales::percent(share, 1)), size = 2, hjust = 2, position = position_stack(vjust = .95)) +
  facet_wrap(~sex) +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(name = "Sex (Reported)", labels = c("Men", "Women"), palette = "Accent") +
  labs(y = "Percent",
       x = NULL,
       title = "Gender Distribution for Each Graduated Major") +
  theme(legend.position = "bottom") +
  scale_y_continuous(expand = c(0,0), labels = scales::percent)

# I would be interested to also include the median salary for each field, but am stuck on how to make that work well visually. Also can't get geom_text() to work here; feedback would be great!

```


## Plot 4
*This is an additional exploration with a different dataset. For my peer reviewer, you can ignore the plot below!*

```{r}

library(albersusa)

us <- usa_sf() %>% 
  rename(state = iso_3166_2)

hate_crime_prop <- hate_crimes %>% 
  rename(post_election = hate_crimes_per_100k_splc,
         pre_election = avg_hatecrimes_per_100k_fbi/6,
         state_full = state,
         state = state_abbrev) %>% 
  select(state, post_election, pre_election) %>%
  pivot_longer(-state,
               names_to = "time_period",
               values_to = "crime_rate")

hate_crime_pre <- hate_crime_prop %>% 
  filter(time_period == "pre_election")

hate_crime_post <- hate_crime_prop %>% 
  filter(time_period == "post_election")

crime_map_before <- left_join(us, hate_crime_pre)

crime_map_after <- left_join(us, hate_crime_post)

#Before
ggplot(us) +
  geom_sf(aes(fill = crime_rate), color = "white", data = crime_map_before) +
  scale_fill_viridis_c(name = "Hate Incidents per 100,000 residents", option = "D") +
  theme_void() +
  theme(legend.position = "bottom")

#After
ggplot(us) +
  geom_sf(aes(fill = crime_rate), color = "white", data = crime_map_after) +
  scale_fill_viridis_c(name = "Hate Incidents per 100,000 residents", option = "D") +
  theme_void() +
  theme(legend.position = "bottom")
```

